---
title: "Tables"
author: "Delsie Johnson"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE)
```

## WNBA Playoffs Table

**Data Source:** \"WNBA_playoffs.csv\" from https://stats.wnba.com/players/boxscores-traditional/. csv file included

**What I want to convey:** I want to show some important game stats for each WNBA team over the course of the 2022 playoffs. These stats include points per game, turnovers per game, and 3 pointers attempted per game. This table can be a good visual for the user to see how each team did.

**Functionality:** I allow the user to highlight each row which makes it easy to see which team they are looking at. There is also a column with graphs of points for each playoff game and the user can hover over the graph and see exactly how many points the team had for each game.

**Formatting:** I put a header over the stats which makes the column names easier to read and understand. I also made each value rounded to one decimal place which is typical of sports stats reporting and makes the table easier to read for the user.

## R Code

```{r}
# loading libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(sparkline)
sparkline(0)

# reading in csv data file
WNBA <- read.csv("WNBA_playoffs.csv", header=T)

# organizing and formatting data
wnba_team <- WNBA %>%
  group_by(team, game_date, w_l) %>% 
  summarise(pts = sum(pts), pa_3 = sum(pa_3), tov = sum(tov))

# creating data frame to use 
playoffs_table <- wnba_team %>% 
  group_by(team) %>% 
  summarise(pts = mean(pts), pa_3 = mean(pa_3), tov = mean(tov)) %>%
  mutate(plt_ppg=NA) %>%
  as.data.frame()

# creating vector for ppg plots
for(t in playoffs_table$team){
  team_ppg <- wnba_team %>%
    filter(team==t) %>%
    arrange(game_date) %>%
    pull(pts)
  playoffs_table[which(playoffs_table$team==t),"plt_ppg"] <- spk_chr(team_ppg)
}

# making table
t_playoffs <- playoffs_table %>%
  select(team, pts, pa_3, tov, plt_ppg) %>%
  kbl(escape = FALSE,
      col.names = c("Team",
                    "Points",
                    "3 Pointers Attempted",
                    "Turnovers",
                    "Points"),
      align = c("l", "r", "r", "r", "r"),
      digits = 1) %>%
  add_header_above(c(" ", "Average Stats per game" = 4)) %>%
  kable_styling(font_size=18)
```

## Display Table in HTML:

```{r display}
t_playoffs %>%
  kable_paper(lightable_options = "hover")
```
